---
title: "Assignment-2"
author: "Ashmi"
date: "January 18, 2018"
output:
  pdf_document: default
  html_document: default
---

#Installing and loading required packages

```{r}
library(tidyverse)
library(ggplot2)
surveys <- read.csv("surveys.csv", header = T, sep = ",") 
View(surveys)
install.packages("gapminder")
library(gapminder)
```
#1. Write R code to extract the survey observations for the first three months of 1990 using the filter() function. (5 points)

```{r}
winter_surveys_90 <- surveys %>% filter((year == "1990") , (month %in% c("1","2","3"))) 
View(winter_surveys)
```

#2. Sort the 1990 winter surveys data by descending order of record ID, then by ascending order of weight.  (10 points)
Taking the winter surveys data as the data of the first three months
```{r}
arrange(winter_surveys_90, desc(record_id))
arrange(winter_surveys_90, weight)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#3. Write code that returns the record_id, sex and weight of all surveyed individuals of Reithrodontomys montanus (RO), (10 points)
```{r}
RO_data <- surveys %>% filter(species_id == "RO") %>% select(record_id, sex, weight)
View(RO_data)
```
#4. Write code that returns the average weight and hindfoot length of Dipodomys merriami (DM) individuals observed in each month (irrespective of the year). Make sure to exclude NA values. (10 points)

```{r}
summarise(group_by(surveys, month, weight_mean = mean(weight, na.rm = T), hindfoot_length_mean = mean(hindfoot_length, na.rm = T)))
```

```{r}

```


