---
title: "Assignment-2"
author: "Ashmi"
date: "January 18, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
yaml:    
  latex-engine: xelatex
---

#Loading required packages
Used supressMessages with library as without it, I was getting this error while knitting: "! Package inputenc Error: Unicode char √ (U+221A) (inputenc) not set up for use with LaTeX."
```{r}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
surveys <- read.csv("surveys.csv", header = T, sep = ",") 
View(surveys)
suppressMessages(library(gapminder))
```
#1. Write R code to extract the survey observations for the first three months of 1990 using the filter() function. (5 points)
```{r}
winter_surveys_90 <- surveys %>% filter((year == "1990") , (month %in% c("1","2","3"))) 
View(winter_surveys_90)
head(winter_surveys_90)
```

#2. Sort the 1990 winter surveys data by descending order of record ID, then by ascending order of weight.  (10 points)
Taking the winter surveys data as the data of the first three months
```{r}
arrange_record <- arrange(winter_surveys_90, desc(record_id), weight)
View(arrange_record)
head(arrange_record)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#3. Write code that returns the record_id, sex and weight of all surveyed individuals of Reithrodontomys montanus (RO), (10 points)
```{r}
RO_data <- surveys %>% filter(species_id == "RO") %>% select(species_id, record_id, sex, weight)
head(RO_data)
```
#4. Write code that returns the average weight and hindfoot length of Dipodomys merriami (DM) individuals observed in each month (irrespective of the year). Make sure to exclude NA values. (10 points)
```{r}
DM <- filter(surveys, species_id =="DM" )
View(DM)
months_DM <- group_by(DM, month)
View(months_DM)
month_sum <- summarise(months_DM, weight_mean = mean(weight, na.rm = T), hindfoot_length_mean = mean(hindfoot_length, na.rm = T))
month_sum
```
#5. Write  code that determines the number of individuals by species observed in the winter of 1990. (15 points)
Taking the winter surveys data as the data of the first three months
```{r}
Groups <- summarize(group_by(winter_surveys_90, species_id), num_individuals =n())
head(Groups)
```
#6. Create a dataframe named gapminder_df and mutate it to contain a column that contains the gross domestic product for each row in the data frame. (5 points)
```{r}
gapminder_df <- as.data.frame(gapminder)
View(gapminder_df)
gapminder_df <- mutate(gapminder_df, GDP = pop*gdpPercap)
head(gapminder_df)
```
#7. Calculate the Mean GDP for Cambodia for the years within the dataset. (15 points)
```{r}
summarize(gapminder_df, country = "Cambodia", mean_GDP = mean(GDP))
```
#8. Find the year with the maximum life expectancy for countries in Asia and arrange them in descending order by year,  The result should contain the country’s name, the year and the life expectancy.  (15 points)
```{r}
gapminder_df %>% filter(continent == "Asia") %>% group_by(country) %>% filter(lifeExp == max(lifeExp)) %>% select(country,year,lifeExp) %>% arrange(desc(year))
```
The year with the maximum life expectancy for countries in Asia is 2007
#9. Count the number of observations per continent. (5 points)
```{r}
gapminder_df %>% group_by(continent) %>% summarise(num_obs=n())
```
#10.  Compute the average and median life expectancy and GDP per capita by continent for the years 1952 and 2007.  Should we be optimistic given the results? (10 points)
```{r}
gapminder_df %>% filter(year %in% c(1952,2007)) %>% group_by(continent, year) %>% summarize(mean_lifeExp = mean(lifeExp), med_lifeExp = median(lifeExp), avg_GDP = mean(gdpPercap), med_GDP = median(gdpPercap))
```
We should be optimistic as the average and median life expectancy and GDP per capita by continent are increasing.
